---
title: "Compare_timepoints"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

load("../data/all_fe_30DPI")
load("../data/all_fe_12week")

all_fe_30DPI
all_fe_12week

all_fe <- rbind(all_fe_12week, all_fe_30DPI)
```

# Visulatizations

Initial identification of populations plot

We first want to view all of the different cell phenotypes within the data
```{r}
# this is the order of markers that we want for all of our plots
order_of_markers <- c("CD3", "CD4", "CD8", "CD45RA", "SLADQ", "CD172")

# to view all of the possible combinations
total_phenotypes <- filter_for_total_pheno(all_fe, marker_order = order_of_markers)

heatmap_all_pheno(total_phenotypes)

# gives the total number of populations
nrow(total_phenotypes) 
```

After identifying all phenotypes, we can filter the data to see the ones that we’re interested in, for example, CD3+ cells that constitute >0.5% of total live leukocytes in a sample.


```{r fig.height=2, fig.width = 1}
# view the specific cell phenotypes we're interested in
sample_populations <- all_fe %>%
  dplyr::filter(CD3 == 1 & percentage > 1) %>%
  arrange(CD3, CD4, CD8) %>%
  filter_pops() 

sample_populations_all_groups <- identified_pop_perc(sample_populations, all_fe, 
                                                     marker_vector = order_of_markers) %>%
  mutate(filename = str_replace(filename, "_Surface_Control_[0-9]*.fcs", ""),
         timepoint = str_extract(filename, "[:alnum:]*_"),
         timepoint = str_replace(timepoint, "_", ""),
         filename = str_extract(filename, "Pig [0-9]*")) %>%
  mutate(vaccine_status = ifelse(filename == "Pig 9947" | filename == "Pig 1817" | filename == "Pig 6059" | filename == "Pig 4515"| filename == "Pig 5273", "vaccinated", "control"))

simple_pop_df <- sample_populations %>%
  column_to_rownames("population") 

simple_pop_df %>%
  dplyr::select(all_of(order_of_markers)) %>%
  mutate_all(~convert_factor_numeric(.)) %>%
    pheatmap::pheatmap(cluster_rows = FALSE, cluster_cols = FALSE,
             labels_row = rownames(simple_pop_df),
             cellwidth = 15, cellheight = 15, angle_col = 45, 
             color = c("#3F4788FF", "#56C667FF"), cutree_rows = 2, legend = FALSE)
```

Correlation Plot
```{r}
corr <- calc_corr(sample_populations_all_groups)

melted_corr <- format_corr(corr)
  
plot_corr(melted_corr) +
    ggplot2::xlab("Populations") +
    ggplot2::ylab("Populations") +
    ggplot2::labs(fill = "Correlation") 
```

Look at percentages
**Note I need to separate out the time and vaccine status (4 groups)
```{r}
ggplot(sample_populations_all_groups, aes(x = timepoint, y = percentage, 
                                   group = vaccine_status, color = vaccine_status)) +
  scale_fill_identity() +
  geom_point() +
  geom_line()
  facet_wrap("population", scales = "free",  labeller = label_both) +
  xlab("Timepoint") +
  ylab("Average Percent of Total Live Leukocytes") +
  theme_gray() 


pops_for_plots_average <- sample_populations_all_groups %>%
  dplyr::group_by(population, vaccine_status) %>%
  dplyr::summarise(average_percent = mean(percentage))

ggplot(pops_for_plots_average, aes(x = vaccine_status, y = average_percent, 
                                   group = vaccine_status, color = vaccine_status)) +
  scale_fill_identity() +
  geom_bar(stat = "identity") +
  facet_wrap("population", scales = "free",  labeller = label_both) +
  xlab("Vaccinate Status") +
  ylab("Average Percent of Total Live Leukocytes") +
  theme_gray() 

```

Anova and Tukey HSD

There are differences between the timepoints

**Note I might actually want to do a paired analysis here...**
```{r}
library(purrr)
library(tidyr)
library(broom)


sample_populations_all_groups %>%
  unite(group, c(timepoint, vaccine_status)) %>%
  mutate(group = as.factor(group)) %>%
  group_by(population) %>%
  nest() %>%
  mutate(aov_result = map(data, ~aov(percentage ~ group, data = .x)),
         tukey_result = map(aov_result, TukeyHSD),
         tidy_tukey = map(tukey_result, tidy)) %>%
  unnest(tidy_tukey, .drop = TRUE) %>%
  dplyr::filter(adj.p.value <0.05) %>%
  select(population, contrast, adj.p.value) %>%
  arrange(paste0("Pop", c(1:length(adj.p.value))))
```

Look at specific phenotypes (double negative, CD4+ CD8+ and double positive

)
```{r}

save_for_phenotypes <- all_fe %>%
  mutate(filename = str_replace(filename, "_Surface_Control_[0-9]*.fcs", ""),
         timepoint = str_extract(filename, "[:alnum:]*_"),
         timepoint = str_replace(timepoint, "_", ""),
         filename = str_extract(filename, "Pig [0-9]*")) %>%
  mutate(vaccine_status = ifelse(filename == "Pig 9947" | filename == "Pig 1817" | filename == "Pig 6059" | filename == "Pig 4515"| filename == "Pig 5273", "vaccinated", "control"))



# look at specific phenotypes (multiple markers)
look_multiple_markers <- function(df, marker, pheno_name) {
  #marker = as.name(marker) # because calling a specific marker
  
  marker_alone <- maker_percent<- NULL # because creating a new column
  
  df %>%
    rename(group = vaccine_status) %>%
    group_by(group, filename, timepoint) %>%
    filter_(marker) %>%
    mutate(marker_alone = sum(cell_no)) %>%
    select(group, filename, timepoint, marker_alone, total_count_by_file) %>%
    unique() %>%
    mutate(marker_percent = 100*marker_alone/total_count_by_file) %>%
  ungroup() %>%
  select(group, filename, timepoint, marker_percent) %>%
    mutate(phenotype = pheno_name)
}

# CD4
pheno_1 <- look_multiple_markers(save_for_phenotypes, c("CD3 == 1 & CD4 == 1"), "CD4_Tcell")


save_for_phenotypes %>%
    rename(group = vaccine_status) %>%
    group_by(group, filename, timepoint) %>%
    filter_(c("CD3 == 1 & CD4 == 1")) %>%
    mutate(phenotype_cells = sum(cell_no)) %>%
    select(group, filename, timepoint, phenotype_cells, total_count_by_file) %>%
    unique() %>%
    mutate(marker_percent = 100*phenotype_cells/total_count_by_file) %>%
  ungroup() %>%
  select(group, filename, timepoint, marker_percent) %>%
    mutate(phenotype = pheno_name)
```
