---
title: "Final PCA Clustering"
author: "Amy Fox"
date: "7/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r message = FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(broom)
library(emmeans)
```

```{r}
# removes columns (metabolites) that have 0 variance across the animals
removeZeroVar <- function(df){
  df[, !sapply(df, function(x) min(x) == max(x))]
}
```

Read in data
```{r}
load("../data/flow_predefined_phenotypes")
load("../data/flow_single_expression")
load("../data/flow_fe")

metabolites <- read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  dplyr::rename(metaboliteID = X) %>%
  select(metaboliteID, contains("Minipig")) %>%
  pivot_longer(cols = c(-metaboliteID), names_to = "full_name", values_to = "metabolite_expression")

metabolite_wide <- metabolites %>%
  filter(!grepl("Pooled", full_name)) %>%
  pivot_wider(names_from = "metaboliteID", values_from = "metabolite_expression") %>%
  removeZeroVar() %>%
  column_to_rownames("full_name")

match_metabolites <- read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  dplyr::rename(metaboliteID = X) 

flow_metab1 <- left_join(flow_predefined_phenotypes, metabolites, by = "full_name") %>%
  rename(percent = marker_percent,
         flow = phenotype,
         vaccine_status = group) 

flow_metab2 <-left_join(flow_fe, metabolites, by = "full_name")%>%
  rename(percent = percentage,
         flow = population)

flow_metab3 <-left_join(flow_single_expression, metabolites, by = "full_name") %>%
  rename(flow = marker)

flow_metab <- rbind(flow_metab1, flow_metab2, flow_metab3) %>%
  mutate(timepoint = str_extract(full_name, "[0-9]*[:alpha:]*$"))

all_flow <- flow_metab %>%
  select(-metaboliteID, -metabolite_expression) %>%
  unique() %>%
  pivot_wider(names_from = flow, values_from = percent)
```

Test variance of metabolites
Test the variance of the metabolites

Normality test - shapiro wilks - if pvalue >0.05, then normal data
Levene's is less sensitive to non-normal data - if pvalue >0.05, then equal variances

If I check the variance across the metabolites, the pvalue <0.05, so non-equal variances
```{r}
library(car)
save <- metabolite_wide %>%
  rownames_to_column("full_name") %>%
  pivot_longer(!full_name, names_to = "metaboliteID", values_to = "metabolite_expression") %>%
  mutate(timepoint = ifelse(grepl("12wks", full_name), "pre", "post")) %>%
  mutate(metaboliteID = as.factor(metaboliteID),
         timepoint = as.factor(timepoint)) %>%
  group_by(metaboliteID) %>%
  nest() %>%
  mutate(norm = map(data, ~shapiro.test(.$metabolite_expression)),
         lev_var = map(data, ~leveneTest(metabolite_expression ~ timepoint, data = .)))%>%
  mutate(tidy_norm = map(norm, broom::glance)) %>%
  unnest(lev_var) %>%
  filter(!is.na(`F value`))

hist(save$`Pr(>F)`)

xx <- save %>%
  rename(p.value = `Pr(>F)`) %>%
  mutate(prop_normal = sum(p.value >0.05)) 

sum(xx$prop_normal)/nrow(xx)

# 60% of the metabolites have normal distribution
# 88% of the metabolites have equal variance according to Levene's test


# OR I want to check variance across metabolites

save <- metabolite_wide %>%
  rownames_to_column("full_name") %>%
  pivot_longer(!full_name, names_to = "metaboliteID", values_to = "metabolite_expression") %>%
  filter(metaboliteID %in% c(sample(1:4483, 500))) %>% # pull out 500 random metab
  mutate(timepoint = ifelse(grepl("12wks", full_name), "pre", "post")) %>%
  mutate(metaboliteID = as.factor(metaboliteID),
         timepoint = as.factor(timepoint)) %>%
  group_by(metaboliteID) %>%
  mutate(var(metabolite_expression),
         sd = sd(metabolite_expression))


  leveneTest(metabolite_expression ~ metaboliteID, data = .)


hist(save$`var(metabolite_expression)`)
hist(save$`sd`)

save %>%
  group_by(metaboliteID) %>%
  mutate(num_0 = sum(metabolite_expression==0))
  select(metaboliteID, sd) %>%
  unique() %>%
  arrange(desc(sd))
  
  # pull out the frist 4 with a lot of var- so doesn't dominate PCA
  # do a PCA for the rest
```

PCA and proportion of variance
```{r}
set.seed(2)
res.pca <- prcomp(metabolite_wide, center = TRUE, scale = TRUE)

res_pca_importance <- summary(res.pca)$importance %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(PC = factor(1:n()))

res_pca_importance %>% 
  ggplot(aes(x = PC, y = `Proportion of Variance`)) + 
  geom_col() + 
  theme_classic()

PCA_minipig <- res.pca$x %>% 
  as.data.frame() %>% 
  rownames_to_column("name") %>%
  mutate(timepoint = str_extract(name, "[:digit:][:digit:][:alpha:]+")) %>% 
  mutate(vaccine_status = ifelse(name == "Minipig_9947_12wks" | name == "Minipig_9947_30DPI"| name == "Minipig_1817_12wks" | name == "Minipig_6059_12wks" | name == "Minipig_4515_12wks"| name == "Minipig_5273_12wks"| name == "Minipig_1817_30DPI" | name == "Minipig_6059_30DPI" | name == "Minipig_4515_30DPI"| name == "Minipig_5273_30DPI", "vaccinated", "control")) %>%
  mutate(timepoint = str_replace(timepoint, "12wks", "pre-infection"),
         timepoint = str_replace(timepoint, "30DPI", "post-infection")) %>%
  mutate(timepoint_vaccine = paste0(timepoint," ", vaccine_status)) 

PCA_minipig%>%
  ggplot(aes(x = PC1, y = PC2, color = timepoint_vaccine)) +
  geom_point() + 
  stat_ellipse() +
  theme_classic() +
  labs(x = paste0("PC1: ", round(res_pca_importance$`Proportion of Variance`[1]*100, 1), "%"),
       y = paste0("PC2: ", round(res_pca_importance$`Proportion of Variance`[2]*100, 1), "%")) +
  ggtitle("Metabolite PCA")


minipig_PCA <- PCA_minipig %>%
  rename(full_name = name) %>%
  mutate(timepoint = str_replace(timepoint,  "pre-infection", "12wks"),
         timepoint = str_replace(timepoint,  "post-infection", "30DPI")) %>%
  left_join(all_flow, by = c("full_name", "timepoint", "vaccine_status")) %>%
  pivot_longer(c(PC1:PC19), names_to = "PC", values_to = "PC_value") %>%
  pivot_longer(c(CD4_Tcell:SLADQ), names_to = "flow", values_to = "flow_percent") %>%
  mutate(timepoint = str_replace(timepoint, "12wks", "pre-infection"),
         timepoint = str_replace(timepoint, "30DPI", "post-infection")) 
```

Interaction model - identify populations where the interaction with timepoint DOES have an effect
```{r}
inter_formula <- flow_percent ~ PC_value *timepoint
flow_percent ~ PC_value + timepoint + PC_value *timepoint
  
# 6 correlations
pops_list <- minipig_PCA %>%
  group_by(PC, flow) %>%
  nest() %>%
  mutate(model = map(data, ~lm(inter_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value) %>%
  filter(term == "PC_value:timepointpre-infection") %>%
  filter(p.value < 0.01) %>%
  arrange(flow, PC) 

pop_list <- minipig_PCA %>%
  group_by(PC, flow) %>%
  nest() %>%
  mutate(model = map(data, ~lm(inter_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value) %>%
  filter(term == "PC_value") %>%
  filter(p.value < 0.01) %>%
  arrange(flow, PC) 

cell <- list(flow_x = pops_list$flow, PC_n = pops_list$PC)

test_trends <- function(df, flow_x, PC_n) {
  
  fm <- df %>%
  filter(flow == flow_x & PC == PC_n) %>%
  mutate(timepoint = recode_factor(timepoint, `1` = "pre-infection", `2` = "post-infection")) %>%
  lm(formula = inter_formula,  data = .)

emtrends(fm, pairwise ~ timepoint, var = "PC_value")$emtrends %>%
  as.data.frame() %>%
  mutate(flow = flow_x,
         PC = PC_n) 
}


# plot the results to see what we're looking at
plot_complex <- function(df, flow_pop, metab_pick) 
  {
fm1 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  lm(formula= inter_formula, data=.)

df3 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  cbind(pred = predict(fm1))
  
ggplot(data=df3, aes(x = PC_value, y = flow_percent, color = factor(timepoint))) + 
  geom_point() + 
  geom_line(mapping=aes(y=pred))  +
  ggtitle(paste(flow_pop, "&", metab_pick)) +
  labs(color = "Timepoint", x = "PC Value", y = paste(flow_pop, "Percentage of Cells"), 
       fill = "Timepoint") 
}

graphs <- pmap(cell, function(flow_x, PC_n) plot_complex(minipig_PCA, flow_x, PC_n))
graphs
library(ggpubr)
do.call(ggarrange, graphs)
```

Additive model - only include correlations not found in interaction model, meaing that the timepoint doesn't have an effect

I could use ggpredict for plotting
https://cran.r-project.org/web/packages/ggiraphExtra/vignettes/ggPredict.html 

Good description of interaction
https://www.medicine.mcgill.ca/epidemiology/joseph/courses/EPIB-621/interaction.pdf
```{r}
add_formula <- flow_percent ~ PC_value + timepoint

# 13 correlations  
pops_list <- minipig_PCA %>%
  group_by(PC, flow) %>%
  filter(flow != "CD8_Tcell" | PC != "PC6") %>%
  filter(flow != "DN_Tcell" | PC != "PC9") %>%
  filter(flow != "Pop12" | PC != "PC2") %>%
  filter(flow != "Pop15" | PC != "PC2") %>%
  filter(flow != "Pop7" | PC != "PC2") %>%
  filter(flow != "Pop8" | PC != "PC2") %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value, estimate) %>%
  filter(term == "PC_value" & p.value < 0.01, .preserve = TRUE) %>%
  arrange(flow, PC)



pops_list2 <- minipig_PCA %>%
    group_by(PC, flow) %>%
  filter(flow == "CD172" & PC == "PC3" |
           flow == "CD45RA" & PC == "PC4" |
           flow == "Pop10" & PC == "PC4" |
           flow == "Pop11" & PC == "PC6" |
           flow == "Pop12" & PC == "PC3" |
           flow == "Pop15" & PC == "PC3" |
           flow == "Pop16" & PC == "PC4" |
           flow == "Pop19" & PC == "PC4" |
           flow == "Pop20" & PC == "PC3" |
           flow == "Pop22" & PC == "PC3" |
           flow == "Pop23" & PC == "PC6" |
           flow == "Pop7" & PC == "PC3" |
           flow == "Pop8" & PC == "PC3") %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value, estimate) %>%
    filter(term =="(Intercept)" & estimate > 0)




# plot the results to see what we're looking at
plot_complex <- function(df, flow_pop, metab_pick) 
  {
fm1 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  lm(add_formula, data = .)

df3 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  cbind(pred = predict(fm1))
  
ggplot(data=df3, aes(x = PC_value, y = flow_percent, color = factor(timepoint))) + 
  geom_point() + 
  geom_line(mapping=aes(y=pred))  +
  ggtitle(paste(flow_pop, "&", metab_pick)) +
  labs(color = "Timepoint", x = "PC Value", y = paste(flow_pop, "Percentage of Cells"), 
       fill = "Timepoint") 
}

cell <- list(flow_x = pops_list$flow, PC_n = pops_list$PC)
pmap(cell, function(flow_x, PC_n) plot_complex(minipig_PCA, flow_x, PC_n))

cell <- list(flow_x = c("CD45RA", "Pop10", "Pop16", "Pop19"), PC_n = rep("PC4", 4))
graphs <- pmap(cell, function(flow_x, PC_n) plot_complex(minipig_PCA, flow_x, PC_n))
do.call(ggarrange, graphs)

p1 <- plot_complex(minipig_PCA, "CD45RA", "PC4")
p2 <- plot_complex(minipig_PCA, "Pop10", "PC4")
p3 <- plot_complex(minipig_PCA, "Pop16", "PC4")
p4 <- plot_complex(minipig_PCA, "Pop19", "PC4")

ggarrange(p1 + rremove("legend") + rremove("xlab"), 
          p2+ rremove("legend") + rremove("xlab"), 
          p3+ rremove("legend"), 
          p4+ rremove("legend"),
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 2, nrow = 2)


pops_list %>%
  select(flow, PC, p.value) %>%
  mutate(p.value = format(p.value, scientific = TRUE, digits = 2)) %>%
  rename(Flow = flow,
         `P value` = p.value)
```
Simple linear model
```{r}
lm_formula <- flow_percent ~ PC_value

# 15 correlations  
pops_list <- minipig_PCA %>%
  group_by(PC, flow) %>%
  filter(flow != "CD8_Tcell" | PC != "PC6") %>%
  filter(flow != "DN_Tcell" | PC != "PC9") %>%
  filter(flow != "Pop12" | PC != "PC2") %>%
  filter(flow != "Pop15" | PC != "PC2") %>%
  filter(flow != "Pop7" | PC != "PC2") %>%
  filter(flow != "Pop8" | PC != "PC2") %>%
  nest() %>%
  mutate(model = map(data, ~lm(lm_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value) %>%
  filter(term == "PC_value") %>%
  filter(p.value < 0.01) %>%
  arrange(flow, PC)

# plot the results to see what we're looking at
plot_complex <- function(df, flow_pop, metab_pick) 
  {
fm1 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  lm(lm_formula, data = .)

df3 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  cbind(pred = predict(fm1))
  
ggplot(data=df3, aes(x = PC_value, y = flow_percent, color = factor(timepoint))) + 
  geom_point() + 
  geom_line(mapping=aes(y=pred), color = "black")  +
  ggtitle(paste(flow_pop, "&", metab_pick)) +
  labs(color = "Timepoint", x = "PC Value", y = paste(flow_pop, "Percentage of Cells"), 
       fill = "Timepoint") 
}

cell <- list(flow_x = pops_list$flow, PC_n = pops_list$PC)
pmap(cell, function(flow_x, PC_n) plot_complex(minipig_PCA, flow_x, PC_n))
```

Interaction and additive

4 output terms
- PC_value:timepoint30DPI
- timepoint30DPI
- PC_value
- (Intercept)

From the 5 correlations picked out
- PC_value - p.value < 0.01 meaning there is a correlation with PC_value
- PC_value:timepoint30DPI - p.value > 0.01 meaning interaction does NOT have an effect
- timepoint30DPI - p.value < 0.01 meaning timepoint has an effect
```{r}
add_inter_formula <- flow_percent ~ PC_value + timepoint + PC_value*timepoint

# 5 correlations  
pops_list <- minipig_PCA %>%
  group_by(PC, flow) %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_inter_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value) %>%
  filter(term == "PC_value" & p.value < 0.01) %>%
  arrange(flow, PC)


minipig_PCA %>%
  group_by(PC, flow) %>%
  filter(flow == "CD45RA" & PC == "PC4" | 
           flow == "Pop10" &  PC == "PC4" |
           flow == "Pop10" & PC == "PC5" | 
           flow == "Pop19" & PC == "PC4" |
           flow == "Pop24" & PC == "PC1") %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_inter_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value)

# plot the results to see what we're looking at
plot_complex <- function(df, flow_pop, metab_pick) 
  {
fm1 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  lm(add_inter_formula, data = .)

df3 <- df %>%
  filter(flow == flow_pop & PC == metab_pick) %>%
  cbind(pred = predict(fm1))
  
ggplot(data=df3, aes(x = PC_value, y = flow_percent, color = factor(timepoint))) + 
  geom_point() + 
  geom_line(mapping=aes(y=pred))  +
  ggtitle(paste(flow_pop, "&", metab_pick)) +
  labs(color = "Timepoint", x = "PC Value", y = paste(flow_pop, "Percentage of Cells"), 
       fill = "Timepoint") 
}

cell <- list(flow_x = pops_list$flow, PC_n = pops_list$PC)
pmap(cell, function(flow_x, PC_n) plot_complex(minipig_PCA, flow_x, PC_n))


```
Pull out PC metabolites

PC4
```{r}
# plot to see where I should cluster
res.pca$rotation %>%
  data.frame() %>%
  ggplot(aes(x = PC4)) +
  geom_histogram(bins = 50) +
  geom_vline(xintercept = 0.03, color = "red") +
  geom_vline(xintercept = -0.03, color = "red") +
  ylab("Weight of Molecular Features")

# 140 metabolites
PC4_cluster1 <- res.pca$rotation %>%
  data.frame() %>%
  filter(PC4 >= 0.03) %>%
  rownames_to_column("metaboliteID") %>%
  mutate(metaboliteID = as.numeric(metaboliteID)) %>%
  select(metaboliteID, PC4)

# 171 metabolites
PC4_cluster2 <- res.pca$rotation %>%
  data.frame() %>%
  filter(PC4 <= -0.03) %>%
  rownames_to_column("metaboliteID") %>%
  mutate(metaboliteID = as.numeric(metaboliteID)) %>%
  select(metaboliteID, PC4)

# left_join(PC4_cluster1, match_metabolites) %>%
#    write.csv("../data/saved_data/PCA/PC4_cluster1_metab_list.csv")
# 
# left_join(PC4_cluster2, match_metabolites) %>%
#    write.csv("../data/saved_data/PCA/PC4_cluster2_metab_list.csv")
```

Look at p-value histograms - these won't look very good because I'm only showing 19 PC on each flow component
```{r}
fit_models <- function(df, markerID) {  
df %>%
  filter(marker == markerID) %>%
  group_by(metaboliteID) %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_formula, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term == "PC_value") %>%
  select(flow, PC, estimate, p.value) %>%
  ungroup()
}
minipig_PCA %>%
  group_by(PC, flow) %>%
  filter(flow != "CD8_Tcell" | PC != "PC6") %>%
  filter(flow != "DN_Tcell" | PC != "PC9") %>%
  filter(flow != "Pop12" | PC != "PC2") %>%
  filter(flow != "Pop15" | PC != "PC2") %>%
  filter(flow != "Pop7" | PC != "PC2") %>%
  filter(flow != "Pop8" | PC != "PC2") %>%
  nest() %>%
  mutate(model = map(data, ~lm(add_formula, data = .)),
         summary_model = map(model, tidy)) %>%
  unnest(summary_model) %>%
  select(flow, PC, term, p.value, estimate)

p_hist <- function(df_fitted_models, markerID) {
  
  alpha = binw = 0.01
pi0 <- df_fitted_models %>%
  mutate(total_rows = nrow(.)) %>%
  filter(p.value >0.5) %>% #
  mutate(pval_0.5 = nrow(.)) %>%
  summarise(pi0 = 2*(pval_0.5/total_rows)) %>%
  unique()
pi1 <- df_fitted_models %>%
  mutate(total_rows = nrow(.)) %>%
  filter(p.value <=alpha ) %>% #
  mutate(pval_alpha = nrow(.)) %>%
  summarise(pi1= (pval_alpha/total_rows)) %>%
  unique()
ggplot(df_fitted_models, aes(x = p.value)) +
  geom_histogram(binwidth = binw) +
  geom_vline(xintercept = alpha, color = "red") +
  geom_hline(yintercept = pi0$pi0*binw * nrow(df_fitted_models), col  = "blue") +
  ggtitle(paste("Metabolite and", markerID, "Percentages P-value Histogram"))
}

fitted_models_SLADQ <- fit_models(minipig_PCA, "SLADQ")
p_hist(fitted_models_SLADQ, "SLADQ")
```

