---
title: "final_identify_metab"
author: "Amy Fox"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(kableExtra)
library(readr)
library(dplyr)
library(purrr)
library(tidyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
```


## Human Metabolome Data Base Search 

1) Go to hmdb.ca 
2) Search, LC-MS
3) Copy mz masses  into “Query Masses 
4) Ion Mode: positive
5) Adduct Type: 

 - M+H
 - M+H-H2O
 - M+NA
 - M+NH4
 
6) Molecular Weight Tolerance: 20 ppm
7) Search
8) Download results as csv

I compiled the metabolites into a single .csv titled "HMDB_PC4_cluster1.csv"/"HMDB_PC4_cluster2.csv"and moved it to the data/saved_data folder in this repo.

Note: We're using the human database because a minpig database does not exist

## MetaboAnalyst - HMDB ID

1) Go to https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml
2) Click on Pathway Analysis 
3) Add in the HMDB ID (Human Metabolome ID) from the "HMDB_PC4_cluster#.csv"
4) Submit


MetaboAnalyst_Results

- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in the selected pathway library
- Homo sapiens (KEGG)

```{r message = FALSE, warning = FALSE}
files <- list.files("../data/saved_data/PCA", pattern = "MetaboAnalyst", full = TRUE)

all_files <- paste0(files, "/pathway_results.csv")

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "MetaboAnalyst_[:graph:]*"))
}

map_df(all_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/pathway_results.csv", "")) %>%
  select(filename, X1, Total, Hits, `Raw p`, Impact) %>%
  mutate(filename = factor(filename, levels = c("PC4_Cluster1", "PC4_Cluster2"))) %>%
  filter(grepl("PC4", filename)) %>%
  arrange(filename) %>%
  rename(p_value = `Raw p`,
         `Putative Pathway` = X1,
         `Correlated Flow` = filename,
         `Metabolites In Pathway` = Total,
         `Identified Metabolites` = Hits) %>%
  mutate(fraction = `Identified Metabolites`/`Metabolites In Pathway`) %>%
  filter(p_value <= 0.01) %>%
  rename(`P value` = p_value) %>%
  kable(align = "l", caption = "Putative Metabolic Pathways Correlated with Flow Cytometry Cell Populations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```


Read in the flow metab data for plotting
```{r}

metabolites <- read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  dplyr::rename(metaboliteID = X) %>%
  select(metaboliteID, contains("Minipig")) %>%
  pivot_longer(cols = c(-metaboliteID), names_to = "full_name", values_to = "metabolite_expression") %>%
  mutate(metaboliteID = str_replace(metaboliteID, "[:digit:]*", paste0("metab_", metaboliteID)))%>%
  filter(!grepl("Pooled", full_name))


load("../data/flow_predefined_phenotypes")
load("../data/flow_single_expression")
load("../data/flow_fe")

flow_metab1 <- left_join(flow_predefined_phenotypes, metabolites, by = "full_name") %>%
  dplyr::rename(percent = marker_percent,
         flow = phenotype,
         vaccine_status = group) 

flow_metab2 <-left_join(flow_fe, metabolites, by = "full_name")%>%
  rename(percent = percentage,
         flow = population)

flow_metab3 <-left_join(flow_single_expression, metabolites, by = "full_name") %>%
  rename(flow = marker)

flow_metab <- rbind(flow_metab1, flow_metab2, flow_metab3) %>%
  mutate(timepoint = str_extract(full_name, "[0-9]*[:alpha:]*$"))%>%
  mutate(timepoint = str_replace(timepoint, "12wks", "pre-infection"),
         timepoint = str_replace(timepoint, "30DPI", "post-infection")) 

unique(flow_metab$flow)
```

```{r}
arrange_plot <- function(df, flow_pop, metab_pick, x_position, yposition1, yposition2) {
  
  step1 <- df %>%
  filter(flow == flow_pop & metaboliteID == metab_pick) %>%
  group_by(flow, metaboliteID) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(flow, metaboliteID, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(flow, metaboliteID, adj.r.squared, p.value, estimate) %>%
  ungroup() %>%
  select(flow, metaboliteID, adj.r.squared, p.value) %>%
  unique() 

# plot after I bring down the number of metabolites to a reasonable number
df %>%
  filter(flow == flow_pop & metaboliteID == metab_pick) %>%
  left_join(step1, by = c("flow", "metaboliteID")) %>%
ggplot(aes(x = metabolite_expression, y = percent, color = factor(timepoint))) +
  geom_point(size = 3) +
  geom_smooth(aes(x = metabolite_expression, y = percent), method = "lm", se = FALSE, color = "#3F4788FF") +
  ggtitle(metab_pick) +
  labs(color = "Timepoint",  fill = "Timepoint") 
}
```


PC4_Cluster2	Arachidonic acid metabolism

PC4: CD45RA, Pop10 (CD3+ CD8+), Pop16 (CD3+ CD4+), Pop19 (CD3+ CD4+ CD45RA+)

Figure for paper: fig.height = 5, fig.width = 4, ncol = 2, nrow = 4
Figure for powerpoint: 
```{r fig.height = 3, fig.width = 4}
hmdb_list <- read.csv("../data/saved_data/PCA/MetaboAnalyst_PC4_Cluster2/name_map.csv") %>%
  filter(KEGG %in% c("C00157", "C00219", "C14768", "C14769", "C14770", "C14771", "C05966", "C05356", "C14823", "C14812", "C14748", "C14749", "C05965", "C14772", "C14773", "C14774", "C14775", "C14781", "C14813", "C04742", "C00909", "C04805", "C14782", "C14814", "C02165", "C02166", "C04853", "C05959", "C00639")) %>%
  select(Query, KEGG) %>%
  rename(compound_id = Query)

cluster_data <- read.csv("../data/saved_data/PCA/HMDB_PC4_cluster2.csv") %>%
  select(-kegg_id) %>%
  filter(compound_id %in% hmdb_list$compound)


mass <- left_join(hmdb_list, cluster_data, by = "compound_id") %>%
  select(compound_id, query_mass, compound_name, KEGG) %>%
  mutate(query_mass = round(query_mass, 5)) %>%
  arrange(query_mass)


read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  mutate(mz = round(mz, 5)) %>%
  filter(mz %in% mass$query_mass)

p1 <- arrange_plot(flow_metab, "CD45RA", "metab_1223",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "CD45RA", "metab_1280",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "CD45RA", "metab_1287",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "CD45RA", "metab_1289",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "CD45RA", "metab_1342",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "CD45RA", "metab_2958",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "CD45RA", "metab_3072",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "CD45RA", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend") + rremove("xylab"), 
          p6+ rremove("legend") + rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 4, nrow = 2) %>%
  annotate_figure(top = text_grob("PC4 Arachadonic acid metabolism and CD45RA", size = 15),
                  left = text_grob("CD45RA Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))



p1 <- arrange_plot(flow_metab, "Pop10", "metab_1223",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop10", "metab_1280",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop10", "metab_1287",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop10", "metab_1289",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop10", "metab_1342",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop10", "metab_2958",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop10", "metab_3072",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop10", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend")+ rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 2, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Arachadonic acid metabolism and Pop10: CD3+ CD8+", size = 15),
                  left = text_grob("Pop10 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))


p1 <- arrange_plot(flow_metab, "Pop16", "metab_1223",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop16", "metab_1280",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop16", "metab_1287",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop16", "metab_1289",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop16", "metab_1342",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop16", "metab_2958",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop16", "metab_3072",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop16", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend")+ rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 2, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Arachadonic acid metabolism and Pop16: CD3+ CD4+", size = 15),
                  left = text_grob("Pop16 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))

p1 <- arrange_plot(flow_metab, "Pop19", "metab_1223",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop19", "metab_1280",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop19", "metab_1287",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop19", "metab_1289",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop19", "metab_1342",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop19", "metab_2958",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop19", "metab_3072",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop19", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 2, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Arachadonic acid metabolism and Pop19: CD3+ CD4+ CD45RA+", size = 15),
                  left = text_grob("Pop19 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))
```

PC4 correlated with PC4: CD45RA, Pop10 (CD3+ CD8+), Pop16 (CD3+ CD4+), Pop24 (CD3+ CD4+ CD8+ CD45RA+)

PC4_Cluster2	Linoleic acid metabolism	

- Phosphatidylcholine
- Linoleate
- 9(10)-EpOME
- 12(13)-EpOME
- (9Z,11E)-(13S)-13-Hydroperoxyoctadeca-9,11-dienoic acid

** PC breakdown is associated with the production of diacylglycerol and the activation of T cells [Aussel 1990]. As we can see on the corresponding correlation plot, as the percentage of CD45RA, a T cell co-receptor, decreases, the expression of the potentially identified phosphatidylcholine  decreases, corroborating the results**
```{r fig.height = 6, fig.width = 7}
hmdb_list <- read.csv("../data/saved_data/PCA/MetaboAnalyst_PC4_Cluster2/name_map.csv") %>%
  filter(KEGG %in% c("C00157", "C01595", "C14825", "C14826", "C04717")) %>%
  select(Query, KEGG) %>%
  rename(compound_id = Query)

cluster_data <- read.csv("../data/saved_data/PCA/HMDB_PC4_cluster2.csv") %>%
  select(-kegg_id) %>%
  filter(compound_id %in% hmdb_list$compound)


mass <- left_join(hmdb_list, cluster_data, by = "compound_id") %>%
  select(compound_id, query_mass, compound_name, KEGG) %>%
  mutate(query_mass = round(query_mass, 5))


read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  mutate(mz = round(mz, 5)) %>%
  filter(mz %in% mass$query_mass)



p1 <- arrange_plot(flow_metab, "CD45RA", "metab_1011",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "CD45RA", "metab_1081",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "CD45RA", "metab_1082",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "CD45RA", "metab_1083",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "CD45RA", "metab_1101",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "CD45RA", "metab_1102",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "CD45RA", "metab_1169",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "CD45RA", "metab_1186",  25, 2E4, 1E4)
p9 <- arrange_plot(flow_metab, "CD45RA", "metab_1263",  25, 2E4, 1E4)
p10 <- arrange_plot(flow_metab, "CD45RA", "metab_1280",  25, 2E4, 1E4)
p11 <- arrange_plot(flow_metab, "CD45RA", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          p9+ rremove("legend")+ rremove("xylab"), 
          p10+ rremove("legend")+ rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 3, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Linoleic acid metabolism and CD45RA", size = 25),
                  left = text_grob("CD45RA Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))


p1 <- arrange_plot(flow_metab, "Pop10", "metab_1011",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop10", "metab_1081",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop10", "metab_1082",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop10", "metab_1083",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop10", "metab_1101",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop10", "metab_1102",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop10", "metab_1169",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop10", "metab_1186",  25, 2E4, 1E4)
p9 <- arrange_plot(flow_metab, "Pop10", "metab_1263",  25, 2E4, 1E4)
p10 <- arrange_plot(flow_metab, "Pop10", "metab_1280",  25, 2E4, 1E4)
p11 <- arrange_plot(flow_metab, "Pop10", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          p9+ rremove("legend")+ rremove("xylab"), 
          p10+ rremove("legend")+ rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 3, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Linoleic acid metabolism and Pop10: CD3+ CD8+", size = 25),
                  left = text_grob("Pop10 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))

p1 <- arrange_plot(flow_metab, "Pop16", "metab_1011",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop16", "metab_1081",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop16", "metab_1082",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop16", "metab_1083",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop16", "metab_1101",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop16", "metab_1102",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop16", "metab_1169",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop16", "metab_1186",  25, 2E4, 1E4)
p9 <- arrange_plot(flow_metab, "Pop16", "metab_1263",  25, 2E4, 1E4)
p10 <- arrange_plot(flow_metab, "Pop16", "metab_1280",  25, 2E4, 1E4)
p11 <- arrange_plot(flow_metab, "Pop16", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          p9+ rremove("legend")+ rremove("xylab"), 
          p10+ rremove("legend")+ rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 3, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Linoleic acid metabolism and Pop16: CD3+ CD4+", size = 25),
                  left = text_grob("Pop16 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))

p1 <- arrange_plot(flow_metab, "Pop19", "metab_1011",  25, 1.5E4, 1E4)
p2 <- arrange_plot(flow_metab, "Pop19", "metab_1081",  25, 2E4, 1E4)
p3 <- arrange_plot(flow_metab, "Pop19", "metab_1082",  25, 2E4, 1E4)
p4 <- arrange_plot(flow_metab, "Pop19", "metab_1083",  25, 2E4, 1E4)
p5 <- arrange_plot(flow_metab, "Pop19", "metab_1101",  25, 2E4, 1E4)
p6 <- arrange_plot(flow_metab, "Pop19", "metab_1102",  25, 2E4, 1E4)
p7 <- arrange_plot(flow_metab, "Pop19", "metab_1169",  25, 2E4, 1E4)
p8 <- arrange_plot(flow_metab, "Pop19", "metab_1186",  25, 2E4, 1E4)
p9 <- arrange_plot(flow_metab, "Pop19", "metab_1263",  25, 2E4, 1E4)
p10 <- arrange_plot(flow_metab, "Pop19", "metab_1280",  25, 2E4, 1E4)
p11 <- arrange_plot(flow_metab, "Pop19", "metab_3602",  25, 2E4, 1E4)

ggarrange(p1 + rremove("legend") + rremove("xylab"), 
          p2+ rremove("legend") + rremove("xylab"), 
          p3+ rremove("legend") + rremove("xylab"), 
          p4+ rremove("legend") + rremove("xylab"), 
          p5+ rremove("legend")+ rremove("xylab"), 
          p6+ rremove("legend")+ rremove("xylab"), 
          p7+ rremove("legend") + rremove("xylab"), 
          p8+ rremove("legend") + rremove("xylab"), 
          p9+ rremove("legend")+ rremove("xylab"), 
          p10+ rremove("legend")+ rremove("xylab"), 
          legend.grob = get_legend(p1), legend = "right", 
          ncol = 3, nrow = 4) %>%
  annotate_figure(top = text_grob("PC4 Linoleic acid metabolism and Pop19: CD3+ CD4+ CD45RA+", size = 25),
                  left = text_grob("Pop19 Percentage of cells", size = 15, rot = 90),
                  bottom = text_grob("Metabolite Expression", size = 15))

```
