---
title: "flow_cleaning_30DPI"
output: html_document
---

Note that I only pulled the "Control" samples because I believe they are unstimulated
*Remove CCR7*

CD8 is different based on the sample
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(data.table)
library(openCyto)
library(ncdfFlow)
library(flowWorkspace)
library(dplyr)
library(ggcyto)
library(stringr) 
library(scales)
library(tidyr)
library(superheat)
library(tibble)
library(pheatmap)
library(cytotypr)
library(gridExtra)
#library(Biobase)
```

Functions for use in the gating

```{r}

.between_interval <- function(x, interval) {
  x <- x[findInterval(x, interval) == 1]
  if (length(x) == 0) {
    x <- NA
  }
  x
}


# this function is from OpenCyto Github RGLab https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R
.find_peaks <- function(x, y = NULL, num_peaks = NULL, adjust = 2, plot = FALSE, ...) {
  x <- as.vector(t(x)) # AMY changed: have to transpose data so that it can find length
  if (length(x) < 2) {
    warning("At least 2 observations must be given in 'x' to find peaks.")
    return(NA)
  }
  if (is.null(y)) {
    dens <- density(x, adjust = adjust, ...)
  } else {
    y <- as.vector(y)
    if (length(x) != length(y)) {
      stop("The lengths of 'x' and 'y' must be equal.")
    }
    dens <- list(x = x, y = y)
  }
  # Discrete analogue to a second derivative applied to the KDE. See details.
  second_deriv <- diff(sign(diff(dens$y)))
  which_maxima <- which(second_deriv == -2) + 1
  # The 'density' function can consider observations outside the observed range.
  # In rare cases, this can actually yield peaks outside this range.  We remove
  # any such peaks.
  which_maxima <- which_maxima[findInterval(dens$x[which_maxima], range(x)) == 1]
  # Next, we sort the peaks in descending order based on the density heights.
  which_maxima <- which_maxima[order(dens$y[which_maxima], decreasing = TRUE)]
  
  # Returns the local maxima. If there are none, we return 'NA' instead.
  if (length(which_maxima) > 0) {
    peaks <- dens$x[which_maxima]
    if (is.null(num_peaks) || num_peaks > length(peaks)) {
      num_peaks <- length(peaks)
    }
    peaks <- peaks[seq_len(num_peaks)]
  } else {
    peaks <- NA
  }
  
  peaks <- data.frame(x = peaks, y = dens$y[which_maxima][seq_len(num_peaks)])
  if(plot){
    plot(dens, main = paste("adjust =" ,  adjust))
    points(peaks, ,col = "red")  
  }
  
  peaks  
}
# this function is from OpenCyto Github RGLab https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R
.find_valleys <- function(x, y = NULL, num_valleys = NULL, adjust = 2, ...) {
  x <- as.vector(t(x))
  if (length(x) < 2) {
    warning("At least 2 observations must be given in 'x' to find valleys.")
    return(NA)
  }
  
  if (is.null(y)) {
    dens <- density(x, adjust = adjust, ...)
  } else {
    y <- as.vector(y)
    if (length(x) != length(y)) {
      stop("The lengths of 'x' and 'y' must be equal.")
    }
    dens <- list(x = x, y = y)
  }
  # Discrete analogue to a second derivative applied to the KDE. See details.
  second_deriv <- diff(sign(diff(dens$y)))
  which_minima <- which(second_deriv == 2) + 1
  # The 'density' function can consider observations outside the observed range.
  # In rare cases, this can actually yield valleys outside this range. We remove
  # any such valleys.
  which_minima <- which_minima[findInterval(dens$x[which_minima], range(x)) == 1]
  # Next, we sort the valleys in descending order based on the density heights.
  which_minima <- which_minima[order(dens$y[which_minima], decreasing = FALSE)]
  # Returns the local minima. If there are none, we return 'NA' instead.
  if (length(which_minima) > 0) {
    valleys <- dens$x[which_minima]
    if (is.null(num_valleys) || num_valleys > length(valleys)) {
      num_valleys <- length(valleys)
    }
    valleys <- valleys[seq_len(num_valleys)]
  } else {
    valleys <- NA
  }
  valleys
}

gate_mindensity_amy <- function(df, channel, filterId = "", positive = TRUE, ignore_values = NULL,
                                marker = NULL,
                       pivot = FALSE, gate_range = NULL, min = NULL, max = NULL,
                       peaks = NULL, ...) {
  
  library(rlang)
  #marker <- as.name(marker)
  
  if (missing(channel) || length(channel) != 1) {
    stop("A single channel must be specified.")
  }
  # Filter out values less than the minimum and above the maximum, if they are
  # given.
  # I altered this to take in a dataframe rather than fcs file
  x <- df[, channel]
  
  if(is.null(ignore_values)) {
    x <- x
  } else{
    x <- x %>%
      filter(CD4 <= ignore_values)
  }
    
  
  if(is.null(peaks))
    peaks <- .find_peaks(x)[, "x"]
  
  if (is.null(gate_range)) {
    gate_range <- c(min(x), max(x))
  } else {
    gate_range <- sort(gate_range)
  }
  
  
  # In the special case that there is only one peak, we are conservative and set
  # the cutpoint as min(x) if 'positive' is TRUE, and max(x) otherwise.
  if (length(peaks) == 1) {
    cutpoint <- ifelse(positive, gate_range[1], gate_range[2])
  } else {
    # The cutpoint is the deepest valley between the two peaks selected. In the
    # case that there are no valleys (i.e., if 'x_between' has an insufficient
    # number of observations), we are conservative and set the cutpoint as the
    # minimum value if 'positive' is TRUE, and the maximum value otherwise.
    valleys <- try(.find_valleys(x), silent = TRUE)
    valleys <- .between_interval(x = valleys, interval = gate_range)
    if (any(is.na(valleys))) {
    #FIXME:currently it is still returning the first peak,
    #we want to pass density instead of x_between to 'min'
    #because x_between is the signal values
      cutpoint <- ifelse(positive, gate_range[1], gate_range[2])
    } else if (length(valleys) == 1) {
      cutpoint <- as.vector(valleys)
    } else if (length(valleys) > 1) {
      # If there are multiple valleys, we determine the deepest valley between
      # the two largest peaks.
      peaks <- sort(peaks[1:2])
      cutpoint <- .between_interval(valleys, peaks)[1]
      # If none of the valleys detected are between the two largest peaks, we
      # select the deepest valley.
      if (is.na(cutpoint)) {
        cutpoint <- valleys[1]
      }      
    }
  }
  gate_coordinates <- ifelse(positive, list(c(cutpoint, Inf)), list(c(-Inf, cutpoint)))
  
  names(gate_coordinates) <- channel
  
  rectangleGate(gate_coordinates, filterId = filterId)
  
  # I added in the print(cutpoint)
  print(cutpoint)
  
}

```


# Pre-infection
Read in the initial gating strategy, which gates to live lymphocytes, into a `gatingTemplate` class.

```{r}
# Identify the file with the gating strategy 
ws <- list.files("../data/", 
                 pattern = "amy_gating_strategy.csv", 
                 full = TRUE)

ws
```

```{r}
# View this template
dtTemplate <- fread(ws)
dtTemplate
```

```{r message = FALSE, warning = FALSE}
# Read in the gating strategy to a 'gatingTemplate' object
initial_gate <- gatingTemplate(ws) 
plot(initial_gate)
```

###  Read in Pre-infection data
```{r}
# Identify the file names of all 20 FCS flow cytometry experiments to 
# read in. 
fcsFiles <- list.files("../data/Surface_D30_DPI", full = TRUE)
fcsFiles


```

```{r}
# Read these files into a 'ncdfFlowSet' object. This will taken a minute
# or so to run. The resulting 'ncdfFlowset' object contains row names with the 
# individual samples and column names with the markers/parameters used in the flow cytometer.
ncfs <- read.ncdfFlowSet(fcsFiles) 
ncfs
```

Apply the initial gating to filter the data to only measurements on live lymphocyte cells.

```{r}
# Convert to a 'GatingSet' object, so we can apply the initial gating to this data.
gs <- GatingSet(ncfs)
gs
```

```{r fig.height = 8, fig.width = 4}
# Apply the initial gating to this data, to filter to only measurements
# on live lymphocyte cells. This may take a minute.
gt_gating(initial_gate, gs)

# You can plot the results of this gating with `autoplot`. For example, to plot
# the gating for the first sample, run:
autoplot(gs[[1]])
```

## Convert data to "tidy data" format

Now that the initial gating has been applied, to limit the data to measurements oflive, singlet lymphocyte cells, we convert the data to a "tidy data" format, to allow us to work with "tidyverse" tools for further analysis and visualization.

```{r}
# Pull out the data from the 'live' node of the gating set (the last node
# in the initial gating strategy).
flowset_gated_data <- gs_pop_get_data(gs, "live") %>% 
  cytoset_to_flowSet() 

# tidy the flowset and convert to a dataframe
df_all_gated <-  tidy_flow_set(flowset_gated_data) %>%
   na.omit() %>%
  dplyr::filter(`SSC-A` != max(`SSC-A`))

unique(df_all_gated$filename)

```

## Clear separation of populations --> Gate with minimum density between + and - populations

We're basically looking at gating 7 different markers

# Try what I did in cyto-fe paper for clinical data

Note I removed CCR7 because there is not a difference in two populations
```{r}
gated_data <- flowset_gated_data

gated_data[[1]]
# initialize the gated data df
all_gated_df = data.frame()

# logicle transform of data in the listed channels 
# convert each of the FMO files to a data frame, adding on the filename
# bind all of the FMO files into one dataframe (though there is only 1 file)
for (i in 1:length(gated_data)) {
  lgcl <- estimateLogicle(gated_data[[i]], channels = c("Alexa Fluor 488-A", "PE-A", "PE-Cy5-A",
                                                        "PE-Cy7-A", "APC-A", "Alexa Fluor 700-A",
                                                        "Alexa Fluor 405-A"))
  
  gated_data_test <- transform(gated_data[[i]], lgcl) %>%
    Biobase::exprs() %>%
    data.frame() %>%
    mutate(filename = sampleNames(gated_data[i])) 
  
  all_gated_df <- rbind(all_gated_df, gated_data_test)
}


all_gated_for_plot <- all_gated_df %>%
  rename("CCR7" = "PE.Cy7.A",
         "SLADQ" = "APC.A",
         "CD3" = "Alexa.Fluor.700.A",
         "CD4" = "PE.Cy5.A",
         "CD8" = "Alexa.Fluor.405.A",
         "CD172" = "Alexa.Fluor.488.A",
         "CD45RA" = "PE.A") %>%
  select(-CCR7)


```

So I think we will need to gate each sample on it's own data (Gate Pig 1817 only on Pig 1817 data) rather than combining all of the samples together 

Calculate the breaks for each sample

```{r}

files <- unique(all_gated_for_plot$filename)

saved_data <- data.frame()
for(i in files) {
  
  pig <- all_gated_for_plot %>%
    filter(filename == i) 
   
  intermediate <- pig %>% 
  mutate(CD3_break = gate_mindensity_amy(., "CD3"),
         CD4_break = gate_mindensity_amy(., "CD4"),
         CD8_break = gate_mindensity_amy(., "CD8"),
         CD45RA_break = gate_mindensity_amy(., "CD45RA"),
         CD172_break = gate_mindensity_amy(., "CD172"),
         SLADQ_break = gate_mindensity_amy(., "SLADQ")) %>%
  select(filename, contains("break")) %>%
  unique() %>%
    print()
  
  saved_data <- rbind(saved_data, intermediate)
}

# I checked for outliers and compared the data to manual gating - in the case of CD172, there was only 1 pig for which the break point was <2, so I found an average of all of the other data points to replace it

# for CD4 and CD45RA, when the break point was >3, there was basically no data that was gated, when we know that there is a clear population separation at a lower range (based on manual gating and just visualization of the populations) - again, I found the average of all of the other values

# I think I should actually find the average of all of the samples and keep the gating consistent

corrected_data <- saved_data %>%
  mutate(CD4_break = ifelse(CD4_break >3, NA, CD4_break),
         CD8_break = ifelse(CD8_break <1, NA, CD8_break),
         CD172_break = ifelse(CD172_break <2, NA, CD172_break),
         CD45RA_break = ifelse(CD45RA_break <1 | CD45RA_break >4, NA, CD45RA_break),
         SLADQ_break = ifelse(SLADQ_break <2.5 | SLADQ_break >3.5, NA, SLADQ_break)) %>%
  mutate(CD3_break = 2.4,
         CD4_break = mean(CD4_break, na.rm = TRUE),
         CD8_break = mean(CD8_break, na.rm = TRUE),
         CD45RA_break = mean(CD45RA_break, na.rm = TRUE),
         CD172_break = 2.2,
         SLADQ_break = mean(SLADQ_break, na.rm = TRUE))
```

Need to check each of the plots now

```{r}

all_data_breaks <- left_join(all_gated_for_plot, corrected_data , by = "filename")

separation_ggplot <- function(df, x_axis, x_title, break_name) {

  ggplot(df, aes(x = x_axis, y = `SSC.A`)) +
   geom_hex(bins = 100) +
   scale_fill_viridis_c() +
   xlab(x_title) +
   ylab("SSC-A") +
    ggtitle(df$filename) +
   geom_vline(xintercept = break_name, size = .7) +
  theme_gray()
}


for(i in files) {
  
  pig_names <- all_data_breaks %>%
    filter(filename == i)
  
  p1 <- separation_ggplot(pig_names, pig_names$CD3, "CD3",
                          pig_names$"CD3_break")
  
  p2 <- separation_ggplot(pig_names, pig_names$CD4, "CD4",
                          pig_names$"CD4_break")
p3 <- separation_ggplot(pig_names, pig_names$CD8, "CD8",
                          pig_names$"CD8_break")
p4 <- separation_ggplot(pig_names, pig_names$CD45RA, "CD45RA",
                          pig_names$"CD45RA_break")
p5 <- separation_ggplot(pig_names, pig_names$CD172, "CD172",
                          pig_names$"CD172_break")
p6 <- separation_ggplot(pig_names, pig_names$SLADQ, "SLADQ",
                          pig_names$"SLADQ_break")


grid.arrange(p1, p2, p3, p4, p5, p6, 
             nrow = 3)
  
}

```

Convert data to fe
```{r}
break_points <- corrected_data %>%
  head(1) %>%
  select(-filename) %>%
  pivot_longer(everything(), names_to = "filename", values_to = "quantile_99") %>%
  mutate(filename = str_replace(filename, "_break", ""))

all_fe_30DPI <- all_data_breaks %>%
  mutate(CD3 = fe(break_points, CD3, "CD3"),
         CD4 = fe(break_points, CD4, "CD4"),
         CD8 = fe(break_points, CD8, "CD8"),
         CD45RA = fe(break_points, CD45RA, "CD45RA"),
         SLADQ = fe(break_points, SLADQ, "SLADQ"),
         CD172 = fe(break_points, CD172, "CD172")) %>%
  select(filename, CD3, CD4, CD8, CD45RA, SLADQ, CD172) %>%
  count_calc()

save(all_fe_30DPI, file = "../data/all_fe_30DPI")
```

# Visulatizations

Initial identification of populations plot

We first want to view all of the different cell phenotypes within the data
```{r}
# this is the order of markers that we want for all of our plots
order_of_markers <- c("CD3", "CD4", "CD8", "CD45RA", "SLADQ", "CD172")

# to view all of the possible combinations
total_phenotypes <- filter_for_total_pheno(all_fe_30DPI, marker_order = order_of_markers)

heatmap_all_pheno(total_phenotypes)

# gives the total number of populations
nrow(total_phenotypes) 
```

After identifying all phenotypes, we can filter the data to see the ones that we’re interested in, for example, CD3+ cells that constitute >0.5% of total live leukocytes in a sample.


```{r}
# view the specific cell phenotypes we're interested in
sample_populations <- all_fe_30DPI %>%
  dplyr::filter(CD3 == 1 & percentage > 1) %>%
  arrange(CD3, CD4, CD8) %>%
  filter_pops() 

sample_populations_all_groups <- identified_pop_perc(sample_populations, all_fe_30DPI, 
                                                     marker_vector = order_of_markers) %>%
  mutate(filename = str_extract(filename, "Pig [0-9]*")) %>%
  mutate(vaccine_status = ifelse(filename == "Pig 9947" | filename == "Pig 1817" | filename == "Pig 6059" | filename == "Pig 4515"| filename == "Pig 5273", "vaccinated", "control"))

simple_pop_df <- sample_populations %>%
  column_to_rownames("population") 

simple_pop_df %>%
  dplyr::select(all_of(order_of_markers)) %>%
  mutate_all(~convert_factor_numeric(.)) %>%
    pheatmap::pheatmap(cluster_rows = FALSE, cluster_cols = FALSE,
             labels_row = rownames(simple_pop_df),
             cellwidth = 15, cellheight = 15, angle_col = 45, 
             color = c("#3F4788FF", "#56C667FF"), cutree_rows = 2, legend = FALSE)
```

Correlation Plot
```{r}
corr <- calc_corr(sample_populations_all_groups)

melted_corr <- format_corr(corr)
  
plot_corr(melted_corr) +
    ggplot2::xlab("Populations") +
    ggplot2::ylab("Populations") +
    ggplot2::labs(fill = "Correlation") 
```

```{r}
ggplot(sample_populations_all_groups, aes(x = vaccine_status, y = percentage, 
                                   group = vaccine_status, color = vaccine_status)) +
  scale_fill_identity() +
  geom_point() +
  facet_wrap("population", scales = "free",  labeller = label_both) +
  xlab("Vaccinate Status") +
  ylab("Average Percent of Total Live Leukocytes") +
  theme_gray() 


pops_for_plots_average <- sample_populations_all_groups %>%
  dplyr::group_by(population, vaccine_status) %>%
  dplyr::summarise(average_percent = mean(percentage))

ggplot(pops_for_plots_average, aes(x = vaccine_status, y = average_percent, 
                                   group = vaccine_status, color = vaccine_status)) +
  scale_fill_identity() +
  geom_bar(stat = "identity") +
  facet_wrap("population", scales = "free",  labeller = label_both) +
  xlab("Vaccinate Status") +
  ylab("Average Percent of Total Live Leukocytes") +
  theme_gray() 

```

T-test
There are no statistically significant differences...
```{r}
library(purrr)
library(tidyr)
library(broom)

sample_populations_all_groups %>%
  group_by(population) %>%
  do(tidy(t.test(percentage ~ vaccine_status, data = .))) %>%
  filter(p.value<0.05)
```

Singular Markers Expression
```{r}
expression <- all_fe_30DPI %>%
  gather(key = marker, value = expression, -filename, -cell_no, -total_count_by_file, -percentage)%>%
  group_by(filename, marker, expression) %>%
  summarise(cells_per_marker = sum(cell_no)) %>%
  ungroup %>%
  group_by(filename, marker) %>%
  mutate(total_cells = sum(cells_per_marker),
         percent_per_marker = 100*cells_per_marker/total_cells) %>%
  ungroup()%>%
  mutate(filename = str_extract(filename, "Pig [0-9]*")) %>%
  mutate(vaccine_status = ifelse(filename == "Pig 9947" | filename == "Pig 1817" | filename == "Pig 6059" | filename == "Pig 4515"| filename == "Pig 5273", "vaccinated", "control"))

expression_plot <- expression %>%
  dplyr::group_by(vaccine_status, marker, expression) %>%
  dplyr::mutate(average_percent_per_marker = mean(percent_per_marker)) %>%
  dplyr::filter(expression == 1) %>%
  select( -cells_per_marker, -total_cells, -percent_per_marker) %>%
  unique()


# all groups together
ggplot(expression_plot, aes(x = vaccine_status,
                                 y = average_percent_per_marker, group = vaccine_status, color = vaccine_status)) +
  scale_fill_identity() +
  geom_point() +
  facet_wrap("marker") +
  xlab("Days Post-Infection") +
  ylab("Average Percent of Cells Expressing Marker") +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 45, size = 13, hjust = 1),
        axis.text.y = element_text(size = 13),
        strip.text.x = element_text(size = 10),
        axis.title = element_text(size = 17),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"))
```



#Check 
```{r}
# check the gating with the percentages
all_data_breaks %>%
  select(filename, CD172_break, CD172) %>%
  group_by(filename) %>%
  mutate(cells = 1) %>%
  mutate(total_cells = sum(cells)) %>%
  filter(CD172>CD172_break) %>%
  mutate(CD172_percentage = 100*sum(cells)/total_cells) %>%
    select(filename, CD172_percentage) %>%
    unique() %>%
  ungroup() %>%
  mutate(average = mean(CD172_percentage))
```

