---
title: "MetaboAnalyst"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Run Correlation with Flow Cytometry

See the "pick_metabolites.Rmd" - then save the .csv titled "Nurul_all_flow_pops.csv"

## Human Metabolome Data Base Search 

1) Go to hmdb.ca 
2) Search, LC-MS
3) Copy mz masses from "Nurul_all_flow_pops.csv" into “Query Masses - note that you can only copy 700 masses at a time, so I did in 2 rounds: rows: 2:500 and then 501:808
4) Ion Mode: positive
5) Adduct Type: 

 - M+H
 - M+H-H2O
 - M+NA
 - M+NH4
 
6) Molecular Weight Tolerance: 20 ppm
7) Search
8) Download results as csv

There were 17,583 compounds identified in HMDB. I compiled the metabolites into a single .csv titled "HMDB_metabolite_search.csv" and moved it to the data/saved_data folder in this repo.

Note: We're using the human database because a minpig database does not exist

## MetaboAnalyst - HMDB ID

1) Go to https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml
2) Click on Pathway Analysis 
3) Add in the HMDB ID (Human Metabolome ID) from the "HMDB_metabolite_search.csv"
4) Submit


MetaboAnalyst_Results1

- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

MetaboAnalyst_Results2
- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Out-degree Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Same top pathways, but differing "Pathway impact", with more pathways with a higher impact and a few more with higher p-value*



Other tests:

Test 3
- Visualization method: Scatter plot
- Enrichment method: Fishers Exact test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Looks like it gives the same results as the MetaboAnalyst_Results1*

Test 4
- Visualization method: Scatter plot
- Enrichment method: Fishers Exact test
- Topology analysis: Out-degree Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Looks like it gives the same results as the MetaboAnalyst_Results2*

Test 5

When I run the "Homo sapiens (SMPDB)" I get a lot of NaN for the results, though some of the pathways pop up such as "Pentose Phosphate Pathway..."

```{r}
library(dplyr)
read.csv("../data/saved_data/MetaboAnalyst_Results1/name_map copy.csv") %>%
  filter(!is.na(Match))

# only 2,907 of the HMDB restuls had a match in the MetaboAnlyst database
```

Tried taking the compound name out of the first result


## MetaboAnalyst - Compound Name

1) Go to https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml
2) Click on Pathway Analysis 
3) Add in the Compound_name from the "HMDB_metabolite_search.csv"
4) Submit

**Looks the same as the 



# Look at the SLADQ-correlated metabolic features

Put the metabolites from "Nurul_all_SLADQ_pops" into HMDB - this technically contains every population that includes SLADQ in my analysis

HMDB_all_SLADQ

MetaboAnalyst_SLADQ

- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

I actually got different results from the overall analysis, and I want to actually correlate the flow with the metabolic pathways, so I need to send each flow population resutls through MetaboAnalyst separately




# Run all of the flow

So Population 6, 9, 13, 14  do not have any results when I put everything through MetaboAnalyst - they have very few features to begin with anyways