---
title: "MetaboAnalyst"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Run Correlation with Flow Cytometry

See the "pick_metabolites.Rmd" - then save the .csv titled "Nurul_all_flow_pops.csv"

## Human Metabolome Data Base Search 

1) Go to hmdb.ca 
2) Search, LC-MS
3) Copy mz masses from "Nurul_all_flow_pops.csv" into “Query Masses - note that you can only copy 700 masses at a time, so I did in 2 rounds: rows: 2:500 and then 501:808
4) Ion Mode: positive
5) Adduct Type: 

 - M+H
 - M+H-H2O
 - M+NA
 - M+NH4
 
6) Molecular Weight Tolerance: 20 ppm
7) Search
8) Download results as csv

There were 17,583 compounds identified in HMDB. I compiled the metabolites into a single .csv titled "HMDB_metabolite_search.csv" and moved it to the data/saved_data folder in this repo.

Note: We're using the human database because a minpig database does not exist

## MetaboAnalyst - HMDB ID

1) Go to https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml
2) Click on Pathway Analysis 
3) Add in the HMDB ID (Human Metabolome ID) from the "HMDB_metabolite_search.csv"
4) Submit


MetaboAnalyst_Results1

- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

MetaboAnalyst_Results2
- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Out-degree Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Same top pathways, but differing "Pathway impact", with more pathways with a higher impact and a few more with higher p-value*



Other tests:

Test 3
- Visualization method: Scatter plot
- Enrichment method: Fishers Exact test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Looks like it gives the same results as the MetaboAnalyst_Results1*

Test 4
- Visualization method: Scatter plot
- Enrichment method: Fishers Exact test
- Topology analysis: Out-degree Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

*Looks like it gives the same results as the MetaboAnalyst_Results2*

Test 5

When I run the "Homo sapiens (SMPDB)" I get a lot of NaN for the results, though some of the pathways pop up such as "Pentose Phosphate Pathway..."

```{r}
library(dplyr)
# read.csv("../data/saved_data/MetaboAnalyst_Results1/name_map copy.csv") %>%
#  filter(!is.na(Match))

# only 2,907 of the HMDB restuls had a match in the MetaboAnlyst database
```

Tried taking the compound name out of the first result


## MetaboAnalyst - Compound Name

1) Go to https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml
2) Click on Pathway Analysis 
3) Add in the Compound_name from the "HMDB_metabolite_search.csv"
4) Submit

**Looks the same as the 



# Look at the SLADQ-correlated metabolic features

Put the metabolites from "Nurul_all_SLADQ_pops" into HMDB - this technically contains every population that includes SLADQ in my analysis

HMDB_all_SLADQ

MetaboAnalyst_SLADQ

- Visualization method: Scatter plot
- Enrichment method: Hypergeometric test
- Topology analysis: Relative-betweeness Centrality
- Reference metabolome: Use all compounds in teh selected pathway library
- Homo sapiens (KEGG)

I actually got different results from the overall analysis, and I want to actually correlate the flow with the metabolic pathways, so I need to send each flow population resutls through MetaboAnalyst separately




# Run all of the flow

So Population 6, 9, 13, 14  do not have any results when I put everything through MetaboAnalyst - they have very few features to begin with anyways

The following seem like they might be good to explore and talk about: CD45RA, CD172, Pop3, Pop 10

Note that these are really the putitive pathways - not actual pathways
```{r message = FALSE, warning = FALSE}
library(readr)
library(kableExtra)
library(purrr) 
library(tidyr)
library(stringr)
library(dplyr)
# all the folders
files <- list.files("../data/saved_data", pattern = "MetaboAnalyst", full = TRUE)

all_files <- paste0(files, "/pathway_results.csv")

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "MetaboAnalyst_[:graph:]*"))
}

map_df(all_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/pathway_results.csv", "")) %>%
  select(filename, X1, Total, Hits, `Raw p`, Impact) %>%
    mutate(filename = factor(filename, levels = c("CD3", "CD4", "CD45RA", "CD172", "SLADQ", 
                                                "DN_Tcell", 
                              "CD4_Tcell", "CD8_Tcell", "Pop1", "Pop3","Pop4","Pop5","Pop6",
                              "Pop7","Pop8","Pop9","Pop10","Pop11","Pop12","Pop13",
                              "Pop14","Pop15","Pop16","Pop17","Pop19","Pop20","Pop21",
                              "Pop22","Pop23","Pop24","Pop25"))) %>%
  arrange(filename) %>%
  rename(p_value = `Raw p`,
         `Putitive Pathway` = X1,
         `Correlated Flow` = filename,
         `Metabolites In Pathway` = Total,
         `Identified Metabolites` = Hits) %>%
  filter(p_value < 0.05) %>%
  rename(`P value` = p_value) %>%
  kable(align = "l", caption = "Putitive Metabolic Pathways Correlated with Flow Cytometry Cell Populations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 



```

# Look at MetaboAnayst for the same-sign metabolites (65 or something of them)

Note: Pop 10 did not have any HMDB references

When I ran CD8 Tcell, CD45RA, Pop8, Pop10 through MetaboAnalyst - didn't get any pathways - could restart computer and recheck

```{r message = FALSE}
files <- list.files("../data/saved_data/Same_sign_metab_analysis", pattern = "MetaboAnalyst", full = TRUE)

all_files <- paste0(files, "/pathway_results.csv")

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "MetaboAnalyst_[:graph:]*"))
}

map_df(all_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/pathway_results.csv", "")) %>%
  select(filename, X1, Total, Hits, `Raw p`, Impact) %>%
  rename(p_value = `Raw p`,
         `Putitive Pathway` = X1,
         `Correlated Flow` = filename,
         `Metabolites In Pathway` = Total,
         `Identified Metabolites` = Hits) %>%
  filter(p_value < 0.05) %>%
  rename(`P value` = p_value) %>%
  kable(align = "l", caption = "Putitive Metabolic Pathways Correlated with Flow Cytometry Cell Populations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 


```

Create table with the following columns where the rows are the flow populations
- # single linear correlations
- # HMDB matches
- # mapped names
- # putative pathways
- # significant putative pathways
```{r message = FALSE}
# single linear corr
single_linear <- read_csv("../data/saved_data/Nurul_metabolites_by_flow_pop.csv") %>%
  select(flow, metaboliteID) %>%
  rename(filename = flow) %>%
  group_by(filename) %>%
  summarise(single_linear_corr = n())

# HMDB matches
files <- list.files("../data/saved_data/HMDB", pattern = ".csv", full = TRUE)

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "HMDB_[:graph:]*"))
}

HMDB_match <- map_df(files, ~read_add_name(.)) %>%
  filter(!grepl("all", filename)) %>%
    filter(!grepl("metabolite_search", filename)) %>%
  mutate(filename = str_replace(filename, "HMDB_", "")) %>%
    mutate(filename = str_replace(filename, ".csv", "")) %>%
  select(compound_id, filename) %>%
  group_by(filename) %>%
  summarise(HMDB_matches = n())


# mapped names
files <- list.files("../data/saved_data", pattern = "MetaboAnalyst", full = TRUE)

name_files <- paste0(files, "/name_map.csv")

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "MetaboAnalyst_[:graph:]*"))
}

mapped_names <- map_df(name_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/name_map.csv", "")) %>%
  select(Comment, filename) %>%
  filter(Comment == 1) %>%
  group_by(filename) %>%
  summarise(mapped_names = n())

# pathways
files <- list.files("../data/saved_data", pattern = "MetaboAnalyst", full = TRUE)

pathway_files <- paste0(files, "/pathway_results.csv")

read_add_name <- function(path) {
  read_csv(path) %>%
    mutate(filename = str_extract(path, "MetaboAnalyst_[:graph:]*"))
}

pathways <- map_df(pathway_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/pathway_results.csv", "")) %>%
  select(X1, filename) %>%
  group_by(filename) %>%
  summarise(putative_pathways = n())

# significant pathways

significant_pathways <- map_df(pathway_files, ~read_add_name(.)) %>%
  filter(!grepl("Results", filename)) %>%
  mutate(filename = str_replace(filename, "MetaboAnalyst_", "")) %>%
  mutate(filename = str_replace(filename, "/pathway_results.csv", "")) %>%
  rename(p_value = `Raw p`) %>%
  filter(p_value < 0.05) %>%
  group_by(filename) %>%
  summarise(significant_pathways = n())

# not sure the best way to order this...
left_join(single_linear, HMDB_match) %>%
  left_join(mapped_names) %>%
  left_join(pathways) %>%
  left_join(significant_pathways) %>%
  mutate_all(~replace_na(., 0)) %>%
  mutate(filename = factor(filename, levels = c("CD3", "CD4", "CD45RA", "CD172", "SLADQ", 
                                                "DN_Tcell", 
                              "CD4_Tcell", "CD8_Tcell", "Pop1", "Pop3","Pop4","Pop5","Pop6",
                              "Pop7","Pop8","Pop9","Pop10","Pop11","Pop12","Pop13",
                              "Pop14","Pop15","Pop16","Pop17","Pop19","Pop20","Pop21",
                              "Pop22","Pop23","Pop24","Pop25"))) %>%
  arrange(filename) %>%
  rename(Flow = filename, 
         `Single Linear Correlation` = single_linear_corr,
         `HMDB Matches` = HMDB_matches,
         `Mapped Names` = mapped_names,
         `Putative Pathways` = putative_pathways,
         `Significant Pathways` = significant_pathways) %>%
  kable(align = "l", caption = "Summary of Molecular Feature and Flow Cytometry Hits") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```







