---
title: "pick_metabolites"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE}
library(data.table)
library(openCyto)
library(ncdfFlow)
library(flowWorkspace)
library(dplyr)
library(ggcyto)
library(stringr) 
library(scales)
library(tidyr)
library(superheat)
library(tibble)
library(pheatmap)
library(cytotypr)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(rstatix)
library(purrr)
library(tidyr)
library(broom)
```

```{r}
removeZeroVar <- function(df){
  df[, !sapply(df, function(x) min(x) == max(x))]
}
```

Metabolite Data
```{r}
metabolites <- read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  dplyr::rename(metaboliteID = X) %>%
  select(metaboliteID, contains("Minipig")) %>%
  pivot_longer(cols = c(-metaboliteID), names_to = "full_name", values_to = "metabolite_expression") %>%
  mutate(metaboliteID = str_replace(metaboliteID, "[:digit:]*", paste0("metab_", metaboliteID)))%>%
  filter(!grepl("Pooled", full_name))


metabolite_wide <- metabolites  %>%
  pivot_wider(names_from = "metaboliteID", values_from = "metabolite_expression") %>%
  mutate(full_name = str_replace(full_name,"12wks", "Week12"),
         full_name = str_replace(full_name,  "30DPI", "D30")) %>%
  removeZeroVar()
         
```

Flow Data
```{r}

load("../data/flow_predefined_phenotypes")
load("../data/flow_single_expression")
load("../data/flow_fe")

flow_metab1 <- left_join(flow_predefined_phenotypes, metabolites, by = "full_name") %>%
  rename(percent = marker_percent,
         flow = phenotype,
         vaccine_status = group) 

flow_metab2 <-left_join(flow_fe, metabolites, by = "full_name")%>%
  rename(percent = percentage,
         flow = population)

flow_metab3 <-left_join(flow_single_expression, metabolites, by = "full_name") %>%
  rename(flow = marker)

flow_metab <- rbind(flow_metab1, flow_metab2, flow_metab3) %>%
  mutate(timepoint = str_extract(full_name, "[0-9]*[:alpha:]*$"))

```

Functions for picking metabolites
```{r}
identify_metab <- function(df, flowID, min_p_value, min_r_squared, timepoint = FALSE) {  
  if(timepoint == FALSE) {
df %>%
  filter(flow == flowID) %>%
  group_by(metaboliteID) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(metaboliteID, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(metaboliteID, adj.r.squared, p.value, estimate) %>%
  ungroup() %>%
  filter(p.value < min_p_value) %>%
  filter(adj.r.squared > min_r_squared) %>%
  select(metaboliteID) %>%
  unique()
    
    } else {
      
    df %>%
  filter(flow == flowID) %>%
  group_by(metaboliteID, timepoint) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(metaboliteID, timepoint, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(metaboliteID, adj.r.squared, estimate, p.value, timepoint) %>%
  ungroup() %>%
  filter(p.value < min_p_value) %>%
  filter(adj.r.squared > min_r_squared) %>%
  select(metaboliteID) %>%
  unique()
  }
}

identify_metab(flow_metab, "SLADQ", 0.01, 0.5)
```

```{r}
identify_metab <- function(df, min_p_value, min_r_squared, timepoint = FALSE) {  
  if(timepoint == FALSE) {
df %>%
  group_by(flow, metaboliteID) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(flow, metaboliteID, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(flow, metaboliteID, adj.r.squared, p.value, estimate) %>%
  ungroup() %>%
  filter(p.value < min_p_value) %>%
  filter(adj.r.squared > min_r_squared) %>%
  select(flow, metaboliteID) %>%
  unique()
    
    } else {
      
    df %>%
  group_by(flow, metaboliteID, timepoint) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(flow, metaboliteID, timepoint, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(metaboliteID, adj.r.squared, estimate, p.value, timepoint) %>%
  ungroup() %>%
  filter(p.value < min_p_value) %>%
  filter(adj.r.squared > min_r_squared) %>%
  select(flow, metaboliteID, timepoint) %>%
  unique()
  }
}

```

# Option 1

This looks at all of the timepoints together and identifies metabolites where 
p <0.01
r^2 >0.5

This will show us the metabolites that are correlated with infection - they may not necessarily be correlated with the flow markers

Took 14 minutes to run - 807 unique metabolies
```{r}

load("../data/saved_data/identity1")
load("../data/saved_data/identity2")
load("../data/saved_data/same_sign_metab")



# 2322 metabolites - (but some may be repeats )
start_time <- Sys.time()

identity1 <- identify_metab(flow_metab,  0.01, 0.5)

identity1 %>%
  select(metaboliteID) %>%
  unique()

save(same_sign_metab, file = "../data/saved_data/same_sign_metab")

end_time <- Sys.time()
end_time - start_time
```

# Option 2
This looks splitting up the timepoints so that there are 2x as many correlations and identifies metabolites where 
p <0.01
r^2 >0.5

There are 3,403 unique metabolites identified
```{r}
# 
start_time <- Sys.time()

identity2 <- identify_metab(flow_metab,  0.01, 0.5, timepoint = TRUE)

identity2 %>%
  select(metaboliteID) %>%
  unique()

end_time <- Sys.time()
end_time - start_time
```

# Option 3
Here we will look at splitting up the timepoints but only taking the metabolites where the r^2 value slope of the line is the same for both of the timepoints. This is a more refined version of "Option 2"
p <0.01
r^2 >0.5

These metabolites show correlation with the flow markers after accounting for the two different timepoints and separating them.

107 unique metabolites
```{r}
start_time <- Sys.time()

fitted_models <- flow_metab %>%
  group_by(flow, metaboliteID, timepoint) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(flow, metaboliteID, timepoint, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(flow, metaboliteID, adj.r.squared, estimate, p.value, timepoint, estimate) %>%
  ungroup() 

end_time <- Sys.time()
end_time - start_time


save_for_plot <-fitted_models %>%
  filter(p.value < 0.01) %>%
  filter(adj.r.squared > 0.5 | adj.r.squared < -0.5) %>%
  select(flow, metaboliteID, timepoint) %>%  
  unique() %>%
  left_join(fitted_models, by = c("metaboliteID", "flow", "timepoint")) %>%
  mutate(same_sign = ifelse(estimate >0, "pos", "neg")) %>%
  group_by(metaboliteID) 



both_neg <- save_for_plot %>%
  filter_at(vars(same_sign), all_vars(. == "neg")) %>%
  group_by(flow) %>%
  mutate(repeate = duplicated(metaboliteID)) %>%
  filter(repeate == "TRUE") %>%
  select(flow, metaboliteID)

both_pos <- save_for_plot %>%
  filter_at(vars(same_sign), all_vars(. == "pos")) %>%
  group_by(flow) %>%
  mutate(repeate = duplicated(metaboliteID)) %>%
  filter(repeate == "TRUE") %>%
  select(flow, metaboliteID)

same_sign_metab <- rbind(both_neg, both_pos) %>%
  left_join(save_for_plot) %>%
  select(-same_sign)

same_sign_metab %>%
  ungroup() %>%
  select(metaboliteID) %>%
  unique()
```


Compare the 3 option lists to each other

Based on this data, 78 of the metabolites that have form parallel lines when correlated with the time separate are also correlated when timepoint isn't accounted for

There are 403 metabolites that correlate both with time and without time
```{r}
#78 overlapping metabolites
same_sign_metab %>%
  ungroup() %>%
  select(metaboliteID) %>%
  inner_join(identity1) %>%
  select(metaboliteID) %>%
  unique()

# all of the metabolites in same_sign are found in identity 2 (that makes sense)
same_sign_metab %>%
  ungroup() %>%
  select(metaboliteID) %>%
  inner_join(identity2) %>%
  select(metaboliteID) %>%
  unique()

# 430 metabolites shared 
inner_join(identity1, identity2) %>%
  select(metaboliteID) %>%
  unique()
```

What I want to do

Option 1 shows us the metabolites that are correlated with infection - they may not necessarily be correlated with the flow markers.
Option 3 (same_sign) shows metabolite correlation with the flow markers after accounting for the two different timepoints and separating them.

goal - have the option 1 (non-time correlated) metabolites - look up - these are metabolites that are associated with infection - if we similarly look at the ones where time is separated and there is still a slope, then (same sign) then these could actually be associated with the flow percentages

```{r fig.width = 7, fig.height = 8}

 pick <- same_sign_metab %>%
  ungroup() %>%
  select(flow, metaboliteID) %>%
  inner_join(identity1) %>%
  select(flow, metaboliteID) %>%
  unique()

unique(pick$flow)

identity1 %>%
  filter(metaboliteID %in% pick$metaboliteID & flow %in% pick$flow)

same_sign_metab  %>%
  mutate(repeate = duplicated(metaboliteID)) %>%
  filter(repeate == "FALSE")



```

See the visualizations for the metabolites and flow that have been identified
```{r}

plot_metab <- function(flowID) {
  
  pick <- same_sign_metab %>%
  ungroup() %>%
  select(flow, metaboliteID) %>%
  inner_join(identity1) %>%
  select(flow, metaboliteID) %>%
  unique() %>%
  filter(flow == flowID)  

flow_metab %>%
  filter(flow == flowID) %>%
  filter(metaboliteID %in% pick$metaboliteID) %>%
ggplot(aes(x = percent, y = metabolite_expression, color = factor(timepoint))) +
  geom_point() +
  geom_smooth(aes(x = percent, y = metabolite_expression), method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~metaboliteID, scales = "free_y") +
  ggtitle(paste("Metabolite Correlations with", flowID, "Percentages"))
}

plot_metab("CD8_Tcell")
plot_metab("Pop8")
plot_metab("Pop10")
plot_metab("Pop4")
plot_metab("Pop9")
plot_metab("CD3")
plot_metab("CD45RA")

```

Show differences when you account for vaccination
```{r}
plot_metab <- function(flowID) {
  
  pick <- same_sign_metab %>%
  ungroup() %>%
  select(flow, metaboliteID) %>%
  inner_join(identity1) %>%
  select(flow, metaboliteID) %>%
  unique() %>%
  filter(flow == flowID)  

flow_metab %>%
  filter(flow == flowID) %>%
  filter(metaboliteID %in% pick$metaboliteID) %>%
  unite(col = "timepoint_vaccine", c(timepoint, vaccine_status), sep = "-") %>%
ggplot(aes(x = percent, y = metabolite_expression, color = factor(timepoint_vaccine))) +
  geom_point() +
  geom_smooth(aes(x = percent, y = metabolite_expression), method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~metaboliteID, scales = "free_y") +
  ggtitle(paste("Metabolite Correlations with", flowID, "Percentages"))
}

plot_metab("CD8_Tcell")
plot_metab("Pop8")
plot_metab("Pop10")
plot_metab("Pop4")
plot_metab("Pop9")
plot_metab("CD3")
plot_metab("CD45RA")
```



```{r}

fit_models <- function(df, markerID, timepoint = FALSE) {  
  if(timepoint == FALSE) {
df %>%
  filter(flow == markerID) %>%
  group_by(metaboliteID) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(metaboliteID, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(metaboliteID, adj.r.squared, p.value, estimate) %>%
  ungroup()
    } else {
      
    df %>%
  filter(flow == markerID) %>%
  group_by(metaboliteID, timepoint) %>%
  nest() %>%
  mutate(model = map(data, ~lm(percent ~ metabolite_expression, data = .)),
         summary_model = map(model, tidy)) %>%
unnest(summary_model) %>%
  filter(term != "(Intercept)") %>%
  select(metaboliteID, timepoint, estimate, model) %>%
  mutate(tidy_model = map(model, broom::glance)) %>%
  unnest(tidy_model) %>%
  select(metaboliteID, adj.r.squared, estimate, p.value, timepoint) %>%
  ungroup() 
  }
}


p_hist <- function(df_fitted_models, markerID) {
  
  alpha = binw = 0.01

pi0 <- df_fitted_models %>%
  mutate(total_rows = nrow(.)) %>%
  filter(p.value >0.5) %>% #
  mutate(pval_0.5 = nrow(.)) %>%
  summarise(pi0 = 2*(pval_0.5/total_rows)) %>%
  unique()

pi1 <- df_fitted_models %>%
  mutate(total_rows = nrow(.)) %>%
  filter(p.value <=alpha ) %>% #
  mutate(pval_alpha = nrow(.)) %>%
  summarise(pi1= (pval_alpha/total_rows)) %>%
  unique()


ggplot(df_fitted_models, aes(x = p.value)) +
  geom_histogram(binwidth = binw) +
  geom_vline(xintercept = alpha, color = "red") +
  geom_hline(yintercept = pi0$pi0*binw * nrow(df_fitted_models), col  = "blue") +
  ggtitle(paste("Metabolite and", markerID, "Percentages P-value Histogram"))
}
```

```{r}
fitted_models_CD8_Tcell <- fit_models(flow_metab, "Pop10")
p_hist(fitted_models_CD8_Tcell, "Pop10")
```

Look at the ones that are in same_sign_meta but not identity1
```{r}

plot_metab <- function(flowID) {
  
  pick <- same_sign_metab %>%
  ungroup() %>%
  select(flow, metaboliteID) %>%
  anti_join(identity1) %>%
  select(flow, metaboliteID) %>%
  unique() %>%
  filter(flow == flowID)  

flow_metab %>%
  filter(flow == flowID) %>%
  filter(metaboliteID %in% pick$metaboliteID) %>%
  unite(col = "timepoint_vaccine", c(timepoint, vaccine_status), sep = "-") %>%
ggplot(aes(x = percent, y = metabolite_expression, color = factor(timepoint_vaccine))) +
  geom_point() +
  geom_smooth(aes(x = percent, y = metabolite_expression), method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~metaboliteID, scales = "free_y") +
  ggtitle(paste("Metabolite Correlations with", flowID, "Percentages"))
}

unique(pick$flow)


plot_metab("CD8_Tcell")
plot_metab("Pop10")
plot_metab("Pop12")
plot_metab("Pop15")
plot_metab("Pop24")
plot_metab("Pop7")
plot_metab("Pop8")
plot_metab("CD3")
plot_metab("DN_Tcell")
plot_metab("Pop4")
plot_metab("Pop9")
```


# Send to Nurul

There are 64 metabolites
```{r}

metabolites <- read.csv("../data/peakTable_Minipig_Vaccinated_Unvaccinated_QC.csv") %>%
  dplyr::rename(metaboliteID = X) %>%
  mutate(metaboliteID = str_replace(metaboliteID, "[:digit:]*", paste0("metab_", metaboliteID)))


same_sign_metab %>%
  ungroup() %>%
  select(flow, metaboliteID) %>%
  inner_join(identity1) %>%
  select(metaboliteID) %>%
  unique() %>%
  left_join(metabolites) %>%
  mutate(metaboliteID = str_replace(metaboliteID, "metab_", "")) %>%
  write.csv("../data/saved_data/metabolites_for_Nurul.csv", row.names = FALSE)
```

